---
title: "LandCover_Analysis"
author: "Anna Wolford"
date: "2024-03-27"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vroom)
library(tidyverse)
library(ggplot2)
library(corrplot)
library(MASS)
library(nlme)
library(leaflet)
library(dplyr)
library(gstat)
library(geoR)
library(multcomp)
source(file="stdres.gls.R")
source(file="predictgls.R")
```

```{r setup1, include=FALSE}
## Read in data
temps <- vroom("SurfaceTemps.txt")
head(temps)
```

## Exploratory Data Analysis

### 1. 
```{r}
ggplot(temps, aes(x = Surface, y = Temp)) +
  geom_boxplot() +
  labs(x = "Surface", y = "Temperature") +
  ggtitle("Temperature Distribution Across Different Surface Types")
```
### 2. 

```{r}
# ggplot(data=temps ,mapping=aes(x=Lon, y=Lat, fill=Temp)) + geom_tile()

# ggplot(data=temps ,mapping=aes(x=Lon, y=Lat, color=Temp)) + geom_point() #or
# ggplot(data=temps ,mapping=aes(x=Lon, y=Lat, fill=Temp)) + geom_raster()

ggplot(data= temps,mapping=aes(x=Lon, y=Lat, color=Temp)) + geom_point() + scale_color_distiller(palette="Spectral",na.value=NA)

ggplot(data= temps,mapping=aes(x=Lon, y=Lat, fill=Temp)) + geom_raster() + scale_fill_distiller(palette="Spectral",na.value=NA)
```
### 3. 

```{r}
temps.lm <- lm(Temp ~ Surface + Lon + Lat, data=temps)
# summary(temps.lm)
# plot(fitted(temps.lm), resid(temps.lm),
#      xlab = "Fitted Values", ylab = "Residuals",
#      main = "Residuals vs. Fitted Values")
# abline(h = 0, col = "red")
temps_clean <- na.omit(temps)
ggplot(data= temps_clean,mapping=aes(x=Lon, y=Lat, fill=resid(temps.lm))) + geom_raster() + scale_fill_distiller(palette="Spectral",na.value=NA)
```

### 4. 

```{r}
temps_clean <- na.omit(temps)
coords <- matrix(c(temps_clean$Lon, temps_clean$Lat), ncol=2, byrow=FALSE)
myVariogram <- variog(coords=coords, data=temps_clean$Temp)
plot(myVariogram)

myVariogram <- variogram(object=Temp~Surface, locations=~Lon+Lat, data=temps_clean)
plot(myVariogram)
```

## Spatial MLR Model Fitting

### 1. Choose with AIC 

```{r cache=TRUE}
model_exp <- gls(Temp ~ Surface, data=temps_clean,
                 correlation=corExp(form=~Lon+Lat, nugget=TRUE), method="ML")

model_spher <- gls(Temp ~ Surface, data=temps_clean,
                   correlation=corSpher(form=~Lon+Lat, nugget=TRUE), method="ML")

model_gaus <- gls(Temp ~ Surface, data=temps_clean,
                  correlation=corGaus(form=~Lon+Lat, nugget=TRUE), method="ML")

summary(model_exp)$AIC
summary(model_spher)$AIC
summary(model_gaus)$AIC ## Gaussian has the lowest AIC value
```
```{r}
coefs <- coef(model_gaus)
cors <- coef(model_gaus$modelStruct$corStruct, unconstrained=FALSE)

coefs
cors

(summary(model_gaus)$sigma)^2
```

## Validating Spatial MLR Model Assumptions and Predictions 

### Linearity 
```{r}
library(car)
car::avPlots(temps.lm)
```

### 2. 

```{r}
sres <- stdres.gls(model_gaus)

# sresm <- matrix(sres, ncol=4, byrow=TRUE)
# round(cor(sresm),2)

residDF <- data.frame(Lon=temps_clean$Lon, Lat=temps_clean$Lat, decorrResid=sres)
residVariogram <- variogram(object=decorrResid~1, locations=~Lon+Lat, data=residDF)
plot(residVariogram)


```

### 3.
```{r}
hist(sres)
```

### 4. 
```{r}

ggplot(data= residDF,mapping=aes(x=Lon, y=Lat, fill=sres)) + geom_raster() + scale_fill_distiller(palette="Spectral",na.value=NA)

# ggplot(data= temps_clean,mapping=aes(x=Lon, y=Lat, fill=resid(temps.lm))) + geom_raster() + scale_fill_distiller(palette="Spectral",na.value=NA)

## Compare
```

### 5. 
```{r}

## CHECK THIS W DR H :) 

system.time({gls(Temp ~ Surface, data=temps_clean,
                  correlation=corGaus(form=~Lon+Lat, nugget=TRUE), method="ML")})
```

```{r}
n.cv <- 50 #Number of CV studies to run
n.test <- nrow(temps_clean)*.2 #Number of observations in a test set
rpmse <- rep(x=NA, times=n.cv)
cvg <- rep(x=NA, times=n.cv)
bias <- rep(x=NA, times=n.cv)
wid <- rep(x=NA, times=n.cv)

n = nrow(temps_clean)
pb <- txtProgressBar(min = 0, max = n.cv, style = 3)
  for(cv in 1:n.cv){
  ## Select test observations
  test.obs <- sample(x=1:n, size=n.test)
  
  ## Split into test and training sets
  test.set <- temps_clean[test.obs,]
  train.set <- temps_clean[-test.obs,]
  
  ## Fit a gls() using the training data ???
  train.lm <- gls(Temp ~ Surface, data=train.set,
                  correlation=corGaus(form=~Lon+Lat, nugget=TRUE), method="ML")
  
  ## Generate predictions for the test set ???
  my.preds <- predictgls(train.lm, newdframe=test.set, level = .95)
  
  ## Calculate RPMSE
  rpmse[cv] <- (test.set[['Temp']]-my.preds[,'Prediction'])^2 %>% mean() %>% sqrt()
  
  ## Calculate Coverage
      cvg[cv] <- ((test.set[['Temp']] > my.preds[,'lwr']) & (test.set[['Temp']] < my.preds[,'upr'])) %>%     mean()
      
        ## Calculate bias
  bias[cv] <- mean(my.preds[,'Prediction']-test.set[['Temp']])
  
  ## Calculate Width
  wid[cv] <- (my.preds[,'upr'] - my.preds[,'lwr']) %>% mean()
  setTxtProgressBar(pb, cv)
    }


close(pb)
```

```{r}
# n.cv <- 50 #Number of CV studies to run
# n.test <- nrow(temps_clean)*.2 #Number of observations in a test set
rpmse.lm <- rep(x=NA, times=n.cv)
cvg.lm  <- rep(x=NA, times=n.cv)
bias.lm  <- rep(x=NA, times=n.cv)
wid.lm <- rep(x=NA, times=n.cv)
 
# n = nrow(temps_clean)

  for(cv in 1:n.cv){
  ## Select test observations
  test.obs <- sample(x=1:n, size=n.test)
  
  ## Split into test and training sets
  test.set <- temps_clean[test.obs,]
  train.set <- temps_clean[-test.obs,]
  
  ## Fit a gls() using the training data ???
  train.lm <- lm(formula=Temp ~ Surface, data=train.set)
  
  ## Generate predictions for the test set ???
  my.preds.lm <- predict.lm(train.lm, newdata=test.set, interval="prediction")
  
  ## Calculate RPMSE
  rpmse.lm[cv] <- (test.set[['Temp']]- my.preds.lm[,'fit'])^2 %>% mean() %>% sqrt()
  
  ## Calculate Coverage
      cvg.lm[cv] <- ((test.set[['Temp']] >  my.preds.lm[,'lwr']) & (test.set[['Temp']] <  my.preds.lm[,'upr'])) %>%     mean()
      
        ## Calculate bias
  bias.lm[cv] <- mean(my.preds.lm[,'fit']-test.set[['Temp']])
  
  ## Calculate Width
  wid.lm[cv] <- (my.preds.lm[,'upr'] - my.preds.lm[,'lwr']) %>% mean()
    }
```


```{r}
mean(rpmse.lm)
  mean(wid.lm)
  mean(cvg.lm)
  mean(bias.lm)
  
  mean(rpmse)
  mean(wid)
  mean(cvg)
  mean(bias)


```

## Statistical Inference

### Use an F-test to see if temperatures are difference across any of the land-cover types.
```{r}
anova(temps.lm)
```

Create confidence intervals for each effect of land cover and determine which land cover types result in increased temperatures.

```{r}
confint(model_gaus) 
```

Only the urban surface results in increased temperatures (with 95% confidence).


Perform a GLHT to construct a confidence interval of the difference temperature between Savannah and Urban land covers.
```{r}
a <- c(1, 0,0,1,0)
b <- c(1,0,0,0,1)
summary_glht <- multcomp::glht(model_gaus, linfct = t(a-b), alternative="two.sided")
confint(summary_glht)
```


Create and map predictions of the temperature at each location that was impeded by cloud cover.

```{r}
temp_nas <- setdiff(temps, temps_clean)
preds.na <- predictgls(model_gaus, newdframe=(temps %>% filter(is.na(Temp))), level=0.95)
preds.na <- preds.na %>%
  mutate(Temp = Prediction)
full_temp <- rbind(preds.na[,1:4], temps_clean)

ggplot(data= full_temp,mapping=aes(x=Lon, y=Lat, fill=Temp)) + geom_raster() + scale_fill_distiller(palette="Spectral",na.value=NA)
```

